---
title: "WWTP"
author: "Emmi Mueller and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

### Overview
Use data from Wu et al. (2019) to test predictions for effects of residence time on biodiversity

#Set up directory
```{r setup, include = FALSE}
rm(list=ls())
require("knitr")
getwd()
```

# Load packages
```{r}
require("png")
require("vegan")
require("ggplot2")
#install.packages("ggpmisc")
#install.packages("cowplot")
require("ggpubr")
require("cowplot")
require("ggpmisc")
```

#Set up figure themes
```{r figure_setup}
my.cols <- RColorBrewer::brewer.pal(n = 4, name = "Greys")[3:4]

# Set theme for figures in the paper
theme_set(theme_classic() + 
  theme(axis.title = element_text(size = 16),
        axis.title.x = element_text(margin = margin(t = 15, b = 15)),
        axis.title.y = element_text(margin = margin(l = 15, r = 15)),
        axis.text = element_text(size = 14),
        axis.text.x = element_text(margin = margin(t = 5)),
        axis.text.y = element_text(margin = margin(r = 5)),
        #axis.line.x = element_line(size = 1),
        #axis.line.y = element_line(size = 1),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.x = element_line(size = 1),
        axis.ticks.y = element_line(size = 1),
        axis.ticks.length = unit(.1, "in"),
        panel.border = element_rect(color = "black", fill = NA, size = 1.5),
        legend.title = element_blank(),
        legend.text = element_text(size = 14),
        strip.text = element_text(size = 14),
        strip.background = element_blank()
        ))
```

# Load data
```{r}
# OTUs (rows) = 96149; sites (columns) = 1188
otu <- read.csv("data/wwtp.otu.csv", sep = ",", header=TRUE)

# Metadata
meta <- read.table("data/wwtp.metadata.txt", sep="\t", header=TRUE)

```


```{r}
#Calculate richness and evenness and add add to metadata
meta$Richness <- NA
for(x in colnames(otu[,-c(1,1188)])){
  meta$Richness[meta$Sample_ID == x] <- specnumber(na.omit(otu[,as.character(x)]))
}

meta$Evenness <- NA
for(x in colnames(otu[,2:1187])){
  meta$Evenness[meta$Sample_ID == x] <- diversity(na.omit(otu[,as.character(x)]), "simpson")
}

#Generate metadata tables for each retention time measurement to remove NAs
metadata_HTR_plant <- subset(meta, meta$HTR_plant != "NA")
metadata_HTR_aeration <- subset(meta, meta$HTR_aeration_tank != "NA")
metadata_SRT <- subset(meta, meta$SRT != "NA")
```


#Code from Wu et al. 2019 
```{r}
hemi<-ifelse(meta$Latitude>0,"North","South")
meta<-data.frame(cbind(meta,hemi=hemi))
ggplot(meta,aes(x=abs(meta$Latitude),y=meta$Richness))+
  geom_point(size=2,alpha=0.6,aes(shape=meta$hemi,colour=meta$SRT))+
  geom_smooth(method = "lm",formula =y ~ poly(x, 2))+
  #scale_colour_manual(values = c("North"="#ff0000","South"="#0000ff"))+
  scale_shape_manual(values = c(19,0))+
  scale_colour_gradientn(colours =c("#aa0000ff","#d40000ff","#ff0000ff","#ff5555ff","#ff8080ff","#ffaaaaff"),trans = 'reverse')+
  xlab("Absolute latitude") +
  ylab("Phylogenetic diversity") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

summary(lm(meta$Richness~I(abs(meta$Latitude))+I(meta$Latitude^2), data=meta))

```

#Regression of Richness ~ Latitude and Residuals ~ SRT and HRT
```{r}
all <- ggplot(meta, aes(y = meta$Richness, x = abs(meta$Latitude)))+
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), col = "red") +
  xlab("abs(latitude") +
  ylab("richness") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_lat_all.pdf")

hrt <- ggplot(metadata_HTR_aeration, aes(y = log(metadata_HTR_aeration$Richness, 10), x = abs(metadata_HTR_aeration$Latitude))) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), col = "red") + 
  xlab("abs(latitude)") + 
  ylab("log(richness)") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_lat_aHRT.pdf")

hrt_model <- lm(log(metadata_HTR_aeration$Richness, 10) ~ poly((abs(metadata_HTR_aeration$Latitude)), 2, raw=TRUE))
metadata_HTR_aeration$Residuals <- hrt_model$residuals

hrt_res <- ggplot(metadata_HTR_aeration, aes(y = metadata_HTR_aeration$Residuals, x = log(metadata_HTR_aeration$HTR_aeration_tank, 10))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ poly(x,2, raw = TRUE), col = "red") +
  xlab("log(aeration HRT)") +
  ylab("Residuals") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_res_aHRT.pdf")

srt <- ggplot(metadata_SRT, aes(y = log(metadata_SRT$Richness, 10), x = abs(metadata_SRT$Latitude))) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), col = "red") + 
  xlab("abs(latitude)") + 
  ylab("log(richness)") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_lat_SRT.pdf")

srt_model <- lm(log(metadata_SRT$Richness, 10) ~ poly((abs(metadata_SRT$Latitude)), 2, raw=TRUE))
metadata_SRT$Residuals <- srt_model$residuals

srt_res <- ggplot(metadata_SRT, aes(y = metadata_SRT$Residuals, x = log(metadata_SRT$SRT, 10))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ poly(x,2, raw = TRUE), col = "red") +
  xlab("log(SRT)") +
  ylab("Residuals") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_res_SRT.pdf")

plot_grid(hrt, hrt_res, srt, srt_res, labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)
ggsave("./output/HRT_SRT.pdf")
```

#Regression for hrt and srt on evenness
```{r}
meta$Inv_Evenness <- NA
for(x in colnames(otu[,2:1187])){
  meta$Inv_Evenness[meta$Sample_ID == x] <- diversity(na.omit(otu[,as.character(x)]), "inv")
}
metadata_HTR_aeration <- subset(meta, meta$HTR_aeration_tank != "NA")

hrt_res <- ggplot(metadata_HTR_aeration, aes(y = metadata_HTR_aeration$Evenness, x = log(metadata_HTR_aeration$HTR_aeration_tank, 10))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ poly(x,2, raw = TRUE), col = "red") +
  xlab("log(aeration HRT)") +
  ylab("Evenness") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_HRT_E.pdf")

hrt_res <- ggplot(metadata_HTR_aeration, aes(y = metadata_HTR_aeration$Inv_Evenness, x = log(metadata_HTR_aeration$HTR_aeration_tank, 10))) +
  geom_point() +
  geom_smooth(method = "lm", formula = y~ poly(x,2, raw = TRUE), col = "red") +
  xlab("log(aeration HRT)") +
  ylab("Evenness") +
  stat_poly_eq(formula = y ~ poly(x,2, raw = TRUE), rr.digits = 2, parse = TRUE)
ggsave("./output/wwtp_HRT_IE.pdf")
```
#Generate figure of richness and evenness with lowess regression for all three RT types 
```{r}
png(filename="output/wwtp_div_rt.png", width = 800, height = 800, res = 96*2)
par(mfrow = c(2,3))

plot(log(metadata_HTR_plant$Richness) ~ log(metadata_HTR_plant$HTR_plant), ylab = "log(S)", xlab = expression(paste("log(", tau, ")")), main = "Plant Hydrologic\nRetention Time")
lines(lowess(log(metadata_HTR_plant$Richness) ~ log(metadata_HTR_plant$HTR_plant)), col = "red")
plot(log(metadata_HTR_aeration$Richness) ~ log(metadata_HTR_aeration$HTR_aeration_tank), ylab = "log(S)", xlab = expression(paste("log(", tau, ")")), main = "Aeration Tank\nHydrologic Retention Time")
lines(lowess(log(metadata_HTR_aeration$Richness) ~ log(metadata_HTR_aeration$HTR_aeration_tank)), col = "red")
plot(log(metadata_SRT$Richness) ~ log(metadata_SRT$SRT), ylab = "log(S)", xlab = expression(paste("log(", tau, ")")), main = "Solid Retention Time")
lines(lowess(log(metadata_SRT$Richness) ~ log(metadata_SRT$SRT)), col = "red")

plot(metadata_HTR_plant$Evenness ~ log(metadata_HTR_plant$HTR_plant), ylab = "E", xlab = expression(paste("log(", tau, ")")))
lines(lowess(metadata_HTR_plant$Evenness~log(metadata_HTR_plant$HTR_plant)), col = "red")
plot(metadata_HTR_aeration$Evenness ~ log(metadata_HTR_aeration$HTR_aeration_tank), ylab = "E", xlab = expression(paste("log(", tau, ")")))
lines(lowess(metadata_HTR_aeration$Evenness~log(metadata_HTR_aeration$HTR_aeration_tank)), col = "red")
plot(metadata_SRT$Evenness ~ log(metadata_SRT$SRT), ylab = "E", xlab = expression(paste("log(", tau, ")")))
lines(lowess(metadata_SRT$Evenness ~ log(metadata_SRT$SRT)), col = "red")

dev.off()
graphics.off()

```
